---
title: "Reproducible Research, Project 1"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

Data loading portion:
```{r chunk 1}

## Load Packages
library(dplyr)
library(ggplot2)

## Load Data
Raw <- read.csv("./activity.csv", header = TRUE)
```

Initial sort of Data
```{r chunk 2}
## Separate Complete and Incomplete Cases
Complete <- Raw[complete.cases(Raw),]
Incomplete <- Raw[!complete.cases(Raw),]
```
For first part of the assignment:
```{r chunk 3}
## Summary of Daily Activity by Date
Daily <- Complete %>% group_by(date) %>% summarize("Activity" = sum(steps))

## Produce Histogram, Add Labels, Title, & Legend 
hist(Daily$Activity, xlab = "Steps", main = "Daily Activity Histogram")
abline(v = mean(Daily$Activity), lty = 1, col = "red", lwd = 3)
abline(v = median(Daily$Activity), lty = 3, col = "blue", lwd = 5)
legend(17000, 20, legend=c("Mean", "Median"), col=c("red", "blue"), lty = 1:3, cex = .8)

## Get Mean and Median of Daily Activity by Date
mean(Daily$Activity)
median(Daily$Activity)
```

Second part of assignment:
```{r chunk 4}
## Summary of Activity by Interval
TimePlot <- Complete %>% group_by(interval) %>% summarize("Avg" = mean(steps))

## Produce Line Plot
plot(TimePlot$interval, TimePlot$Avg, type = "l", xlab = "Interval", ylab = "Activity Avg.", main = "Time Plot of Activity")

## Find Interval with Highest Activity
TimePlot[which.max(TimePlot$Avg), 1]
```

Imputing Data and recreating graphs (note I've used mean of initial data)
For unknown reason, knitr is not able to produce mean / median lines as RStudio does (see png file).
```{r chunk 5}
## Impute data with mean of steps from initial data set & create a new data frame 
Incomplete$steps[is.na(Incomplete$steps)] = mean(Raw$steps, n.rm = TRUE)
ImpMerge <- rbind(Incomplete, Complete)

## Summarize Imputed Data
ImpDaily <- ImpMerge %>% group_by(date) %>% summarize("Activity" = sum(steps))
```
```{r Chunk 5 Graph}
## Histogram of Imputed Data
hist(ImpDaily$Activity, xlab = "Steps", main = "Daily Activity Histogram - Imputed w/ Mean # of Steps")
abline(v = mean(ImpDaily$Activity), lty = 1, col = "red", lwd = 3)
abline(v = median(ImpDaily$Activity), lty = 3, col = "blue", lwd = 5)
legend(17000, 20, legend=c("Mean", "Median"), col=c("red", "blue"), lty = 1:3, cex = .8)
```
 
Adding days and classifying as Weekday or Weekend
```{r chunk 6}
## Add Weekday Info & Classify as Weekday or Weekend
ImpMerge$Day <- weekdays(as.Date(ImpMerge$date, format = "%Y-%m-%d"))
ImpMerge$DayClass <- ifelse(ImpMerge$Day == "Saturday" | ImpMerge$Day == "Sunday", "Weekend", "Weekday")
```

Last part of assignment:
```{r chunk 7}
## Summarize Modified Data
ImpTimePlot <- ImpMerge %>% group_by(interval, DayClass) %>% summarize("Avg" = mean(steps))
```

Plot for Activity Weekend vs. Weekday is produced in R Studio (also attached as .png), but is not able to print when knitr is used to create html file.
```{r}
a <- ggplot(ImpTimePlot, aes(interval, Avg, color = DayClass))
a + geom_line() + ggtitle("Activity Time Plot, Weekday v Weekend") + xlab("Intervals") + ylab("Activity")
```
